### 2022-02-24 paper/referee_report_1/response.md
Update and rename README.md to response.md

1. The observation, theory and scientific application of pulsars are not fully introduced and need to be expanded. For example, This, along with tests of general relativity (e.g. Kramer et al. 2006; Miao et al. 2021, ApJ, 921, 114), pulsar timing arrays for gravitational wave detection (e.g. Manchester et al. 2013), pulsar braking and magnetospheric dynamics (e.g. Gao et al. 2017, 849,19, Wang et al. 2020, Universe, 6, 63) the neutron star equation of state (e.g. Antoniadis et al. 2013 Science 340 448, Pang, et al. 2021, ApJ, 922,14), and other pulsar science applications, motivates efforts to find new pulsars via large scale pulsar surveys, which continue to be conducted up to the present day.

  * We have added in the suggested text and references.

2. There are some peculiar but interesting subpulse drifting pulsars, such as J1727−2739， J1822-2256 and the Vela pulsar, which could be missed in this paper. Please refer to articles: Wen et al. 2016, A&A 592, A127; Wen et al. 2021,ApJ,900,168; Janagal, Parul et al. 2022, MNRAS, 509, 4573.

  * We have added the text "J0026−1956 is also found to contain null sequences, a phenomenon which is known to be connected to drifting behaviour, originally noticed by (Lyne & Ashworth 1983), and further attested by recent studies of PSRs J1727−2739 (Wen et al. 2016), J1822−2256 (Janagal et al. 2022), and the Vela pulsar (Wen et al. 2020)" to the introduction.

3. The authors referred to the RS model (Ruderman & Sutherland 1975). Although the RS model successfully explains the formation mechanism of drift subpulse, it is difficult to explain the complex drift subpulse phenomena, such as the erratic drift rate, the change of drift direction, and the transition between various drift states. For this reason, many researchers have developed the RS model to explain various special subpulse drifting phenomena (e.g. Gil & Sendyk 2000,ApJ, 541,351; Qiao et al. 2004, ApJ,616, L127).

  * We have edited the text to read "Such pulsars exist, and several extensions have been proposed over the years to account for them. For example, Gil & Sendyk (2000) suggest that a quasi-central spark can account for (non-drifting) core components in profiles, and the well-known phenomenon of bi-drifting may be explained by presence of an inner annular gap (Qiao et al. 2004), an inner acceleration gap (Basu et al. 2020), or non-circular spark motions (Wright & Weltevrede 2017). Such extensions are typically developed to explain specific drifting behaviors that are observed in a relatively small subset of pulsars, and there still lacks a single, comprehensive theory that can describe all drifting behaviors. [newline] Pulsars that both show complicated drifting behavior and which are bright enough for single pulse analysis, are relatively rare."

4. Which bright source was used for phase or amplitude calibration? Is it possible to convert the flux into absolute amplitude in the figures instead of arbitrary unit?

  * On inspection, it turns out that all of our observations were calibrated on 3C444. I have edited the text to read "Each observation was individually calibrated using observations of 3C444 taken within ∼2 hours of the respective target observations. The calibration solutions were obtained using the Real Time System software (Mitchell et al. 2008)."
  * Getting reliable absolute flux calibration is challenging for the MWA because of its large, complex primary beam. A fuller investigation of the absolute brightness of the pulsar is intended for a follow-up paper, for which we have also recently acquired observations at higher frequencies (uGMRT), and which we will use for spectral analysis.

5. Please clarify the technical principle of the PSRCHIVE's pdmp to determine period and DM.

6. In single-pulse study, the radio frequency interference can act to reduce the signal-to-noise ratio. Considering the weak nature of single pulses, it is necessary to subtract away or mask undesirable signals from the data before single-pulse analysis. Please describe the RFI mitigation technique you used.

  * We have added the text: "The MWA is located in a very radio quiet location, and no significant RFI was found in the beamformed data."

7. Please indicate the flux value in the ordinates of Figure 1 for each averaged pulse profiles.

  * Since we do not have reliable absolute fluxes at this stage, Figure 1 shows normalised profiles. We have added "peak-normalised" to the caption of the figure.

8. It is difficult to determine whether nulling or just weak emission exists in this pulsar only from Figures 2 and 3. The pulse energy distribution is a statistical way to characterize the nulling behavior (Wen et al. 2020, ApJ, 904, 72).

9. Is there any temporal evolution of nulling fraction from follow-up observations and archival MWA data? Is it possible that the constraints on the nulling fraction might be reported in this paper?

10. The nulls from a few pulsars does not seem to be random, but show well-defined periodicity (e.g. Herfindal & Rankin 2007, MNRAS, 380, 430; Wang et al. 2021, ApJ, 923, 259). Try to examine whether the periodicity in the transitions between null and burst states is present in this pulsar.

11. The pulsar switches between null and burst states with short and long timescales, which provides an important clue to scrutinize various nulling mechanism models. The null and burst length histograms are required.

12. Whether the observed nulls are true or just weak emission below the detection threshold of the telescope? To answer this question, the analysis of integrated pulse profile obtained from all classified nulls is necessary.

13. There was no detection in the two 20-min MWA observations made via periodicity search. A few pulsars were reported to exhibit sporadic, strong single pulses coexisting with a periodic weak emission in the duration of weak mode (e.g. PSR B0826-34). Does J0026-1956 has the similar emission behavior? You'd better try single-pulse search technique on the two observations.

  * We formed pulsestacks of the two 20-min observations and found a small amount of relatively faint emission in them that is just visible in the pulse stack. Accordingly, we have promoted these two observations to the same standing as the original four, included them in Table 1 and Figure 2, and performed all the same analyses on them. The lack of significant emission in these two observations means that the main conclusions of our nulling analysis are essentially unchanged. There are many places throughout the text that have been updated accordingly, which are marked in bold as usual.

14. The LRFS in Figure 6 clearly presents the drifting feature at a frequency of ~0.03 cycles/period in the on-pulse window. In addition, a wide modulation feature with frequency ranging from 0.05-0.1 cycles/period is detected in the leading emission component. The corresponding feature with weak power in the 2DFS is also discernible.

15. The polarized pulse profiles for different drift modes provide an important clue to reveal the triggering mechanism of mode changing. Is there any difference between averaged pulse profiles by integrating pulses for different drift modes?

16. The observed different subpulse drift modes and the slowly changing in drift rate may be an indication of a local change in the electromagnetic field configuration in the gap. A lot of theoretical modes have been proposed, such as van Leeuwen et al. 2003, A&A, 399, 223; Smith et al. 2005, A&A, 440, 683; Yuen, 2019, MNRAS, 486, 2011.

17. To our knowledge, no model can satisfactorily explain the observed subpulse drift characteristics for different pulsars. In the future, the multi-frequency study will be required.